---
output:
  md_document:
    variant: markdown_github
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

**NOTE: This is a toy package created for expository purposes. It is not meant to actually be useful. If you want a package for factor handling, please see [forcats](https://cran.r-project.org/package=forcats).**

### foofactors

Factors are a very useful type of variable in R, but they can also drive you nuts. This package provides some helper functions for the care and feeding of factors.

### Installation

```{r installation, eval = FALSE}
devtools::install_github("qiaoyuet/foofactors")
```
  
### Quick demo

#### fbind()

Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

#### freq_out()

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work. Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```


#### detect_fct()

R is not so good as detecting whether a factor is actually just characters. We consider a factor is a true factor when the number of unique values in it does not equal to the length of it. For example, this is a factor with two levels:

```{r}
a <- factor(c("A","A","B"))
length(unique(a))
length(a)
detect_fct(a)
```

If the number of unique values does equal to the length of the factor we consider it as characters instead of a factor. For example, this should be characters instead of a factor with three levels:

```{r}
b <- factor(c("A","B","C"))
length(unique(b))
length(b)
detect_fct(b)
```

#### reorder_desc()

This function can reorder the levels of a factor into descending order. The descending levels of strings are automatically considered as descending alphabetical order in R.

```{r}
a <- factor(c("1","3","2","6"))
levels(a)
levels(reorder_desc(a))

b <- factor(c("Statistics","Mathematics","Computer Science"))
levels(b)
levels(reorder_desc(b))
```

#### factor.asis()

This function allows you to set levels as the unique values that appear in the data. For example, we want the levels to be "high","low","medium" instead of "high","low","low","high","medium".

```{r}
a <- factor(c("high","low","low","high","medium"))
factor_asis(a)
```

